The Migration of the Field Code

### 2016-02-17 talks/dh-working-group/dh-working-group.Rmd
Corrections from Kellen and Abby and additional citations

After the American Revolution, most states were common law jurisdictions, sometimes with courts of chancery. These courts had a complex system of pleading defined by case law. By the 1840s, lawyers and the mercantile classes called for the simplification and rationalization of civil procedure through codification. Factions for and against codification debated about whether codes were laws passed by democratic legislatures or anti-democratic legal elites, whether codification served only the needs of wealthy capitalists, and whether the purity of Anglo-Saxon civilization derived from the common law would be maintained. The economic capital New York was the first state to codify its procedure in 1848, thanks to the efforts of David Dudly Field. By the end of the century, New York's Field Code became the model for the codes of civil procedure in most states.

@@

Legal historians have long known that the Field code spread to other jurisdictions. Beyond the mere fact of its adoption, however, no one has studied the content of the borrowings. Which codes borrowed from each other? Which sections were borrowed, and howe were they modified? What were the patterns and structures of borrowings and of innovations? To answer these questions we gathered a corpus of 115 codes and statutes of civil procedure from 1806 to 1933 containing about 7.6 million words, then algorithmically detected the borrowings within that corpus.

@@

We found out how the codes borrowed from one another by splitting the codes into sections and comparing each section to every other section.^[The data and code to re-run our analyses are available in a GitHub repository: <https://github.com/lmullen/civil-procedure-codes>. We generalized our method in Lincoln Mullen, "textreuse: Detect Text Reuse and Document Similarity," R package version 0.1.2 (2015): <https://github.com/ropensci/textreuse>, which was peer-reviewed by rOpenSci.] This process in essence mimicked the way that nineteenth-century code commissioners literally cut and pasted sections from other jurisidictions.

@@

Having identified all of the relevant laws of civil procedure in the nineteenth-century, including codes, session laws, and statutes, we used OCR software to create plain-text versions of the codes. These OCR files received only a light cleaning: we edited the section markers by hand as necessary, and wrote a script to fix the most obvious OCR errors.

We then split each section of each code into its own text file. The corpus contains nearly 98,000 sections. Below is a sample file containing a single section from a single code.^[1851 California Laws 74, §151. This corresponds to our file `CA1851-001660.txt`. Notice the minor OCR errors.]

@@

Next we used the Jaccard similarity score for measuring document similarity. That measure, treating the tokens as a set, is the ratio of shared tokens to the ratio of total tokens in two documents. That ration can range from 0 (complete dissimilarity) to 1 (complete similarity).^[Jure Leskovec, Anand Rajaraman, and Jeff Ullman, *Mining of Massive Datasets*, 2nd ed. (Cambridge University Press, 2014), ch. 3, <http://www.mmds.org>. The formal definition of the Jaccard similarity score is $$J(A, B) = \frac{ | A \cap B| }{ | A \cup B|}$$] For instance, the section from the 1851 California code above was derived from the 1850 New York code.^[*Final Report of the Commission on Practice and Pleadings*, 2 *Documents of the Assembly of New York*, 73rd session, No. 16 (1850), 317, §756. This corresponds to our file `NY1850-008350.txt`.] Because of changes to the wording and OCR errors, the Jaccard similarity between the two sections was `r jaccard_similarity(sections[["NY1850-008350"]], sections[["CA1851-001660"]]) %>% round(3)`. By carefully checking matching sections versus scores, we arrived at a rule of thumb that a Jaccard similarity score greater than 0.15 likely indicated a match, and a score greater than 0.2 almost certainly indicated a match.^[The mean score for the best section matches was `r best_matches$score %>% mean(na.rm = TRUE) %>% round(2)`.]

@@

This matrix required filtering based on what we knew about the process of borrowing.\marginnote{Read this table as \texttt{row} borrows from \texttt{column}.} For instance, a code from 1851 obviously did not borrow from a code from 1877. Furthermore, in chains of borrowing (e.g., NY1850 $\rightarrow$ CA1851 $\rightarrow$ CA1868 $\rightarrow$ CA1872 $\rightarrow$ MT1895) the latest section might have a high similarity to all of its parents, but it was in fact borrowed only from the most recent parent. We therefore filtered the similarity matrix to remove (1) matches within the same code; (2) anachronistic matches; (3) spurious matches beneath a certain threshold. Then if a section had multiple matches, we kept the match from (5) the chronologically closest code, (6) giving preference to codes from the same state, unless (7) there was a substantially higher match from a different code. The result was a sparse matrix of the most likely matches for each section, with at most one match for each section.

@@

The result was a set of approximately 2,900 clusters which contained at least five sections, though this probably overstates the number of ur-sections in the corpus. Each cluster contained a list of the section that belonged to it. The biggest cluster, for instance, which concerned the use of affidavits in pleading, contained 103 sections. Scholars in digital literary studies often speak of the "deformance" of texts.^[Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (University of Illinois Press, 2011), ch. 3.] Clustering the sections of the codes deformed them by pulling the sections out of their context within specific codes and by arranging them by topic in chronological order. Deforming the codes allowed us to see the development and spread of the law. Consider this (admittedly more significant than typical) excerpt from a cluster of sections about the competence of witnesses:

@@

By framing our historical questions around "medium data" we enjoyed a useful symbiosis of traditional and digital historical methods. Our computational methods produced useful historical knowledge they were justified by what we knew about the data from traditional historical work. We knew that code commissioners worked with "the scissors and paste-pot,"^[*Rocky Mountain News*, January 20, 1877.] as political debates from the era frequently complained, and we knew codes in the archives that commisions literally marked up the legislation of other states.^[The Nevada code was created by marking up a printed copy of California's code. Detail from Council Bill 21, First Territorial Legislative Session (1861), Nevada State Library, Archives and Public Records.] Knowing that borrowings happened section by section justified our use of the minhash/LSH algorithm, and the nature of the borrowings justified our use of affinity propagration clustering. And knowing how codes passed from state to state made network analysis an obvious fit.^[This is an example of the "no free lunch theorem" in action. See David Robinson, "K-means clustering is not a free lunch," *Variance Explained*, January 16, 2015: <http://varianceexplained.org/r/kmeans-free-lunch/>, citing David H. Wolpert and William G. Macready, "No Free Lunch Theorems for Optimization" *IEEE Transactions on Evolutionary Computation* 1, no. 1 (April 1997): 67--82, <http://ti.arc.nasa.gov/m/profile/dhw/papers/78.pdf>.]

@@

We are writing an article with a fuller statement of our historical interpretations, but to sum up what we learned from the computational methods: This study is significant because it gets at the heart of lawmaking in U.S. history. Lawyers and judges, politicians and newspaper editors warred over whether codes that were drafted by commissioners and borrowed wholesale from one another were actually democratic laws or the imposition of a legal elite. Furthermore, these borrowings called into question the extent to which the United States was actually a federal system of laws. We have shown the dominance of New York's laws in a supposedly federal system, and demonstrated the extent to which code commissioners borrowed from the Field Code and its descendants. Despite the notion that states were equal and soverign, economic centers were in fact legal centers. But we have also demonstrated that regional distinctions were also significant, with states like California and Ohio creating separate Western and Midwestern families of codes. Finally we have shown how haphazard the process of codification could be, as in the case of the 1855 Washington code cited above, or in the case of New Mexico, which borrowed Missouri's laws because those were the law books available to the commissioners. In most jurisdictions the exact language of the Field Code is not currently on the books, but its basic provisions for civil procedure are in force throughout the United States.^[But in some states such as Missouri the Field Code provisions survive nearly unchanged.] Without too much exaggeration we might say that our method has revealed the spine of modern American legal practice.

>>>>

After the American Revolution, most states were common law jurisdictions, sometimes with courts of chancery. These courts had a complex system of pleading defined by case law. By the 1840s, lawyers and the mercantile classes called for the simplification and rationalization of civil procedure through codification. Factions for and against codification debated about whether codes were laws passed by democratic legislatures or anti-democratic legal elites, whether codification served only the needs of wealthy capitalists, and whether the purity of Anglo-Saxon civilization derived from the common law would be maintained. The economic capital New York was the first state to codify its procedure in 1848, thanks to the efforts of David Dudley Field. By the end of the century, New York's Field Code became the model for the codes of civil procedure in most states.

@@

Legal historians have long known that the Field Code spread to other jurisdictions. Beyond the mere fact of its adoption, however, no one has studied the content of the borrowings. Which codes borrowed from each other? Which sections were borrowed, and how were they modified? What were the patterns and structures of borrowings and of innovations? To answer these questions we gathered a corpus of 115 codes and statutes of civil procedure from 1806 to 1933 containing about 7.6 million words, then algorithmically detected the borrowings within that corpus.

@@

We found out how the codes borrowed from one another by splitting the codes into sections and comparing each section to every other section.^[The data and code to re-run our analyses are available in a GitHub repository: <https://github.com/lmullen/civil-procedure-codes>. We generalized our method in Lincoln Mullen, "textreuse: Detect Text Reuse and Document Similarity," R package version 0.1.2 (2015): <https://github.com/ropensci/textreuse>, which was peer-reviewed by rOpenSci. We benefitted a great deal from  David A. Smith, Ryan Cordell, and Elizabeth Maddock Dillon, "Infectious Texts: Modeling Text Reuse in Nineteenth-Century Newspapers," in *2013 IEEE International Conference on Big Data*, 2013, 86–94, doi:10.1109/BigData.2013.6691675; David A. Smith et al., "Detecting and Modeling Local Text Reuse," in *Proceedings of the 14th ACM/IEEE-CS Joint Conference on Digital Libraries* (IEEE Press, 2014), 183–92, <http://dl.acm.org/citation.cfm?id=2740800>.] This process in essence mimicked the way that nineteenth-century code commissioners literally cut and pasted sections from other jurisidictions.

@@

Having identified all of the relevant laws of civil procedure in the nineteenth century, including codes, session laws, and statutory compilations, we used OCR software to create plain-text versions of the codes. These OCR files received only a light cleaning: we edited the section markers by hand as necessary, and wrote a script to fix the most obvious OCR errors.

We then split each section of each code into its own text file.^[While all statutes are divided into chapters, articles, and titles, specific regulations appear within these as sequentially numbered sections.] The corpus contains nearly 98,000 sections. Below is a sample file containing a single section from a single code.^[1851 California Laws 74, §151. This corresponds to our file `CA1851-001660.txt`. Notice the minor OCR errors.]

@@

Next we used the Jaccard similarity score for measuring document similarity. That measure, treating the tokens as a set, is the ratio of shared tokens to the ratio of total tokens in two documents. That ratio can range from 0 (complete dissimilarity) to 1 (complete similarity).^[Jure Leskovec, Anand Rajaraman, and Jeff Ullman, *Mining of Massive Datasets*, 2nd ed. (Cambridge University Press, 2014), ch. 3, <http://www.mmds.org>. The formal definition of the Jaccard similarity score is $$J(A, B) = \frac{ | A \cap B| }{ | A \cup B|}$$] For instance, the section from the 1851 California code above was derived from the 1850 New York code.^[*Final Report of the Commission on Practice and Pleadings*, 2 *Documents of the Assembly of New York*, 73rd session, No. 16 (1850), 317, §756. This corresponds to our file `NY1850-008350.txt`.] Because of changes to the wording and OCR errors, the Jaccard similarity between the two sections was `r jaccard_similarity(sections[["NY1850-008350"]], sections[["CA1851-001660"]]) %>% round(3)`. By carefully checking matching sections versus scores, we arrived at a rule of thumb that a Jaccard similarity score greater than 0.15 likely indicated a match, and a score greater than 0.2 almost certainly indicated a match.^[The mean score for the best section matches was `r best_matches$score %>% mean(na.rm = TRUE) %>% round(2)`.]

@@

This matrix required filtering based on what we knew about the process of borrowing.\marginnote{Read this table as \texttt{row} borrows from \texttt{column}, e.g., \texttt{CA1851-004380} borrows from \texttt{NY1850-018680}.} For instance, a code from 1851 obviously did not borrow from a code from 1877. Furthermore, in chains of borrowing (e.g., NY1850 $\rightarrow$ CA1851 $\rightarrow$ CA1868 $\rightarrow$ CA1872 $\rightarrow$ MT1895) the latest section might have a high similarity to all of its parents, but it was in fact borrowed only from the most recent parent. We therefore filtered the similarity matrix to remove matches within the same code; anachronistic matches; spurious matches beneath a certain threshold. Then if a section had multiple matches, we kept the match from the chronologically closest code, giving preference to codes from the same state, unless there was a substantially higher match from a different code. The result was a sparse matrix of the most likely matches for each section, with at most one match for each section.

@@

The result was a set of approximately 2,900 clusters which contained at least five sections, though this probably overstates the number of ur-sections in the corpus. Each cluster contained a list of the section that belonged to it. The biggest cluster, for instance, which concerned the use of affidavits in pleading, contained 103 sections. Scholars in digital literary studies often speak of the "deformance" of texts.^[Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (University of Illinois Press, 2011), ch. 3.] Clustering the sections of the codes deformed them by pulling the sections out of their context within specific codes and by arranging them by topic in chronological order. Deforming the codes allowed us to see the development and spread of the law. Consider this excerpt from a cluster of sections about the competence of witnesses:

@@

By framing our historical questions around "medium data" we enjoyed a useful symbiosis of traditional and digital historical methods. Our computational methods produced useful historical knowledge because they were justified by what we knew about the data from traditional historical work. We knew that code commissioners worked with "the scissors and paste-pot,"^[*Rocky Mountain News*, January 20, 1877.] as political debates from the era frequently complained, and we knew of codes in the archives that commisions literally marked up the legislation of other states.^[The Nevada code was created by marking up a printed copy of California's code. Detail from Council Bill 21, First Territorial Legislative Session (1861), Nevada State Library, Archives and Public Records.] Knowing that borrowings happened section by section justified our use of the minhash/LSH algorithm, and the nature of the borrowings justified our use of affinity propagration clustering. And knowing how codes passed from state to state made network analysis an obvious fit.^[This is an example of the "no free lunch theorem" in action. See David Robinson, "K-means clustering is not a free lunch," *Variance Explained*, January 16, 2015: <http://varianceexplained.org/r/kmeans-free-lunch/>, citing David H. Wolpert and William G. Macready, "No Free Lunch Theorems for Optimization" *IEEE Transactions on Evolutionary Computation* 1, no. 1 (April 1997): 67--82, <http://ti.arc.nasa.gov/m/profile/dhw/papers/78.pdf>.]

@@

We are writing an article with a fuller statement of our historical interpretations, but to sum up what we learned from the computational methods: This study is significant because it gets at the heart of lawmaking in U.S. history. Lawyers and judges, politicians and newspaper editors warred over whether codes that were drafted by commissioners and borrowed wholesale from one another were actually democratic laws or the imposition of a legal elite. Furthermore, these borrowings called into question the extent to which the United States was actually a federal system of laws. We have shown the dominance of New York's laws in a supposedly federal system, and demonstrated the extent to which code commissioners borrowed from the Field Code and its descendants. Despite the notion that states were equal and sovereign, economic centers were in fact legal centers. But we have also demonstrated that regional distinctions were also significant, with states like California and Ohio creating separate Western and Midwestern families of codes. Finally we have shown how haphazard the process of codification could be, as in the case of the 1855 Washington code cited above, or in the case of New Mexico, which borrowed Missouri's laws because those were the law books available to the commissioners. In most jurisdictions the exact language of the Field Code is not currently on the books, but its basic provisions for civil procedure are in force throughout the United States.^[But in some states such as Missouri the Field Code provisions survive nearly unchanged.] Without too much exaggeration we might say that our method has revealed the spine of modern American legal practice.
